---
title: "My report"
format:
  docx:
    fig-cap-location: top
    reference-doc: data/template.docx
execute:
  echo: false # R code is not printed
  message: false # R messages are not printed
  warning: false # R warning messages are not printed
---

```{r, eval=FALSE}
# make sure you have the latest version of the packages
#install.packages("pak")
pak::pkg_install(c("visOmopResults", "omopgenerics", "IncidencePrevalence", "CohortCharacteristics"))
```

```{r}
# Check needed packages ----
if (packageVersion("visOmopResults") < "1.5.1") {
  stop("visOmopResults >= 1.5.1 is required")
}
if (packageVersion("omopgenerics") < "1.4.2") {
  stop("omopgenerics >= 1.4.2 is required")
}

# Load necessary packages ----
library(visOmopResults)
library(here)
library(readr)
library(omopgenerics)
library(dplyr)
library(IncidencePrevalence)
library(CohortCharacteristics)

# Load results ----
results <- importSummarisedResult(path = here("rawData"))

# source functions
source(file.path(getwd(), "functions.R"))

# Global options ----
knitr::opts_chunk$set(
  out.width  = "95%",  # figures occupy ~95% of document width
  out.height = "auto",
  dpi        = 320,    # ensure figure quality
  fig.width  = 6,      # default aspect ratio (can be overridden per-figure)
  fig.height = 3,
  results    = "asis"  # enable Markdown produced via cat() inside chunks
)

setGlobalPlotOptions(type = "ggplot") # the best configuration for word documents
setGlobalTableOptions(type = "flextable") # the best configuration for word documents
```

```{r}
# read the new names and sections for the report
dic <- read_csv("data/dictionary.csv", show_col_types = FALSE)
```

```{r}
# you can define some functions here so you can reuse them
formatCount <- function(x) {
  format(round(x), big.mark = ",")
}
formatPercentage <- function(x, n = 2) { # by default formated to 2 decimals
  sprintf(paste0("%.", n, "f%%"), x)
}
```

Let's first process the data to change the names of the cohorts to the provided results. In the characterisation results the cohort name is stored in group (cohort_name) whereas for the prevalence results the names are stored in group (outcome_cohort_name) the first thing we will do is to split the results into prevalence and characterisation and then change the results.

```{r}
resultsCharacterisation <- results |>
  filterResult(result_type == "summarise_characteristics") |>
  splitGroup() |>
  left_join(
    dic |>
      select("cohort_name" = "original_name", "new" = "report_name"),
    by = c("cohort_name")
  ) |>
  select(!"cohort_name") |>
  rename("cohort_name" = "new") |>
  uniteGroup(cols = "cohort_name") |>
  newSummarisedResult()

resultsPrevalence <- results |>
  filterResult(result_type == "prevalence") |>
  splitGroup() |>
  left_join(
    dic |>
      select("outcome_cohort_name" = "original_name", "new" = "report_name"),
    by = c("outcome_cohort_name")
  ) |>
  select(!"outcome_cohort_name") |>
  rename("outcome_cohort_name" = "new") |>
  uniteGroup(cols = c("denominator_cohort_name", "outcome_cohort_name")) |>
  newSummarisedResult()
```

Does the code make sense?

In this first block of exercises we will work only with the neurological conditions.

```{r}
neuroCond <- dic |>
  filter(section == "Neurological conditions") |>
  pull("report_name")
```

You can use the variable that we have just created to print the neurological conditions programmatically. There are `r length(neuroCond)` neurological conditions, they are: `r neuroCond`.

## Simple paragraph

Let's try to make our PI happy we have agreed that we want to include the following text:

The total number of individuals under observation was XXX.

The neurological condition with the highest prevalence was XXX, with a prevalence of XX% (95% CI: XX–XX%), while the condition with the lowest prevalence was XXXX, with a prevalence of XX% (95% CI: XX–XX%).

The median age at diagnosis ranged from XX years (IQR: XX–XX) for XXXX condition to XX years (IQR: XX–XX) for XXXX condition.

The proportion of females ranged from XX% in XXXX condition to XX% in XXXX condition.

When generating the text programmatically, consider:

- How many decimal places should be reported? You can use/define functions in the block that we created above.
- What should happen when two or more conditions have the same minimum or maximum value?
- How should conditions with no observed cases be handled?

To start with do this with only database 1:

```{r}
xc <- resultsCharacterisation |>
  filterResult(
    cohort_name %in% neuroCond,
    cdm_name == "Database 1",
    variable_name %in% c("Age", "Sex"),
    estimate_name %in% c("q25", "median", "q75", "percentage")
  ) |>
  tidy()
xp <- resultsPrevalence |>
  mutate(variable_name = "prevalence") |>
  filterResult(
    outcome_cohort_name %in% neuroCond,
    cdm_name == "Database 1"
  ) |>
  tidy() |>
  select(
    "cdm_name", "outcome_cohort_name", "variable_name", "outcome_count", 
    "denominator_count", "prevalence", "prevalence_95CI_lower",
    "prevalence_95CI_upper"
  )
```

To help you get started you the first sentence could be done like:

```{r}
numObs <- unique(xp$denominator_count)
```

The total number of individuals under observation was `r formatCount(numObs)`.

...


## Simple Table

Now we will do a table to show the prevalence of the different conditions of interest in the different databases use `tablePrevalence()` after filtering to the conditions of interest. Populate the arguments of tablePrevalence so the table looks nice.

```{r}
resultsPrevalence |>
  filterResult() |> # populate the filter accordingly
  tablePrevalence( # populate header, groupColumn, hide to make the table as you wish
    type = "flextable",
    style = "_brand.yml"
  )
```

Note that for this prevalence table we had to explicitly state the style and type, this is not needed for the rest of the packages as they automatically detect the style and type from global options. In the future (hopefully) this will be implemented and you won't need to manually state the style and type.

Now we want to add a caption to this table to do so we can use the predefined style template, see the example below:

:::{custom-style="TableCaption"}
Table 1. This is the title for table 1.
:::

```{r}
emptyTable()
```

:::{custom-style="CaptionFooter"}
IQR = interquartile range
:::

Use this example to create your own table with caption.

## Simple Figure

Similarly you can create a figure plot the Prevalence of the different conditions of interest (the same than before) using the plotPrevalence function, customise the plot so it looks as you wish.

```{r}
resultsPrevalence |>
  filterResult() |> # populate the filter accordingly
  plotPrevalence( # populate x, y, colour, facet, ... to make the table as you wish
  )
```

## Loop of Sections

Now it would be nice to repeat the paragraph, the figure and the table that you created in a loop for all the different sections. When you have multiple outputs the "game" changes a bit. The first thing that you need to know is that you must use the option results = "asis" in the configuration of the r chunk. You will also need to use helper functions that have been defined in `functions.R`.

Let's see how it will look:

```{r, results='asis'}
sections <- unique(dic$section)
for (sec in sections) {
  # conditions of interest
  conditions <- dic$report_name[dic$section == sec]
  
  # to create the section title
  cat(paste0("### ", sec, "\n\n"))
  
  # data of interest
  xc <- resultsCharacterisation |>
    filterResult(
      cohort_name %in% conditions,
      variable_name %in% c("Age", "Sex"),
      estimate_name %in% c("q25", "median", "q75", "percentage")
    ) |>
    tidy()
  xp <- resultsPrevalence |>
    mutate(variable_name = "prevalence") |>
    filterResult(outcome_cohort_name %in% conditions) |>
    tidy() |>
    select(
      "cdm_name", "outcome_cohort_name", "variable_name", "outcome_count", 
      "denominator_count", "prevalence", "prevalence_95CI_lower",
      "prevalence_95CI_upper"
    )
  
  # Try to fit this text:
  # The total number of individuals under observation was XXX.
  # The neurological condition with the highest prevalence was XXX, with a
  # prevalence of XX% (95% CI: XX–XX%), while the condition with the lowest
  # prevalence was XXXX, with a prevalence of XX% (95% CI: XX–XX%).
  # The median age at diagnosis ranged from XX years (IQR: XX–XX) for XXXX
  # condition to XX years (IQR: XX–XX) for XXXX condition.
  # The proportion of females ranged from XX% in XXXX condition to XX% in XXXX
  # condition.
  
  # For example:
  formatProportion <- function(x) {
    sprintf("%.2f%% for %s in %s", x$percentage, x$cohort_name, x$cdm_name)
  }
  xfm <- xc |>
    filter(variable_level == "Female") |>
    filter(percentage == min(percentage))
  xfM <- xc |>
    filter(variable_level == "Female") |>
    filter(percentage == max(percentage))
  propFem <- paste0(
    "The proportion of females ranged from ",
    formatProportion(xfm),
    " to ",
    formatProportion(xfM),
    ".\n\n"
  )
  cat(propFem)
  
  # create tables
  captionTable("Table XX. Title")
  
  emptyTable() |>
    printTable() # note that now explicit printing is needed
  
  footer("Footer")
  
  # create figures
  
  emptyPlot() |>
    printFigure() # note that now explicit printing is needed
  
  captionFigure("Figure XX. Title")
  
  footer("Footer")
  
}
```

Note that when you try to generalise, several problems will arise, sections that have just one conditions, conditions that only have males in the denominator, conditions that are not present in a certain database, counts under 5, ... This will test you code to the limit, good luck :)

## Spacing in quarto

Probably
you
have
noticed
that adding just one
line break
does
not add the line break in the rendered document. If you want to 
create a new line you need

two line breaks.

Also you can add `{{< pagebreak >}}` if you want to add

{{< pagebreak >}}

a page break. As you can see multiple line






breaks do not add extra line breaks (e.g. space/1 line break -> space; 2 or more line breaks -> 1 line break).

If you explicitly want to add a line break you must use `\` see example:

\

This added two line breaks.

## Change the style

- The style of tables and figures is controlled by the `_brand.yml` file change the darwin blue definition to: '#AF2F3E' or change the white lines into black or play with it. Regenerate the document so you can check that the changes have been applied properly. 

- The style of captions and text is controlled by the `data/template.docx` see on the styles part that *TableCaption*, *FigureCaption*, *CaptionFooter* have been defined. Play with them and add for example bold to the FigureCaption or change the colour or whatever you like. Regenerate the document so you see that the changes have been affected.

- Use the functions that you defined before to define the number of decimals used to display a percentage.

- Change the number of decimals used in the tables, you can control that using the `options` argument.

- For titles of different levels use '#' the more that you use will determine which is the level of the style used: e.g. if you use #### (4 hashtags) this will match the Heading 4 style.
