---
title: "My report"
format:
  docx:
    fig-cap-location: top
    reference-doc: data/template.docx
execute:
  echo: false # R code is not printed
  message: false # R messages are not printed
  warning: false # R warning messages are not printed
---

```{r, eval=FALSE}
# make sure you have the latest version of the packages
if (!requireNamespace("pak", quietly = TRUE)) install.packages("pak")
pak::pkg_install(c(
  "visOmopResults", "omopgenerics", "IncidencePrevalence",
  "CohortCharacteristics", "here", "readr", "dplyr"
))
```

```{r}
# Check needed packages ----
if (packageVersion("visOmopResults") < "1.5.1") {
  stop("visOmopResults >= 1.5.1 is required")
}
if (packageVersion("omopgenerics") < "1.4.2") {
  stop("omopgenerics >= 1.4.2 is required")
}

# Load necessary packages ----
library(visOmopResults)
library(here)
library(readr)
library(omopgenerics)
library(dplyr)
library(IncidencePrevalence)
library(CohortCharacteristics)

# Load results ----
results <- importSummarisedResult(path = here("rawData"))

# source functions
source(file.path(getwd(), "functions.R"))

# Global options ----
knitr::opts_chunk$set(
  out.width  = "95%",  # figures occupy ~95% of document width
  out.height = "auto",
  dpi        = 320,    # ensure figure quality
  fig.width  = 6,      # default aspect ratio (can be overridden per-figure)
  fig.height = 3,
  results    = "asis"  # enable Markdown produced via cat() inside chunks
)

setGlobalPlotOptions(type = "ggplot") # the best configuration for word documents
setGlobalTableOptions(type = "flextable") # the best configuration for word documents
```

```{r}
# read the new names and sections for the report
dic <- read_csv("data/dictionary.csv", show_col_types = FALSE)
```

```{r}
# you can define some functions here so you can reuse them
formatCount <- function(x) {
  format(round(x), big.mark = ",")
}
formatPercentage <- function(x, n = 2) { # by default formated to 2 decimals
  sprintf(paste0("%.", n, "f%%"), x)
}
```

First, we will update the cohort names using the supplied dictionary. In the characterisation results, the cohort name is stored in the group field `cohort_name`, whereas in the prevalence results it is stored in `outcome_cohort_name`. We will therefore split the results into characterisation and prevalence results before renaming the cohorts.

```{r}
resultsCharacterisation <- results |>
  filterResult(result_type == "summarise_characteristics") |>
  splitGroup() |>
  left_join(
    dic |>
      select("cohort_name" = "original_name", "new" = "report_name"),
    by = c("cohort_name")
  ) |>
  select(!"cohort_name") |>
  rename("cohort_name" = "new") |>
  uniteGroup(cols = "cohort_name") |>
  newSummarisedResult()

resultsPrevalence <- results |>
  filterResult(result_type == "prevalence") |>
  splitGroup() |>
  left_join(
    dic |>
      select("outcome_cohort_name" = "original_name", "new" = "report_name"),
    by = c("outcome_cohort_name")
  ) |>
  select(!"outcome_cohort_name") |>
  rename("outcome_cohort_name" = "new") |>
  uniteGroup(cols = c("denominator_cohort_name", "outcome_cohort_name")) |>
  newSummarisedResult()
```

Does the code make sense?

In this first block of exercises we will work only with the neurological conditions.

```{r}
neuroCond <- dic |>
  filter(section == "Neurological conditions") |>
  pull("report_name")
```

You can use the variable that we have just created to print the neurological conditions programmatically. There are `r length(neuroCond)` neurological conditions: `r neuroCond`.

## Simple paragraph

Let's make our PI happy. We have agreed to include the following text:

The total number of individuals under observation was XXX.

The neurological condition with the highest prevalence was XXX, with a prevalence of XX% (95% CI: XX–XX%), while the condition with the lowest prevalence was XXXX, with a prevalence of XX% (95% CI: XX–XX%).

The median age at diagnosis ranged from XX years (IQR: XX–XX) for XXXX condition to XX years (IQR: XX–XX) for XXXX condition.

The proportion of females ranged from XX% in XXXX condition to XX% in XXXX condition.

When generating the text programmatically, consider:

- How many decimal places should be reported? You can use/define functions in the block that we created above.
- Prevalence and its confidence limits are stored as proportions (for example, `0.0431`), so multiply them by 100 before passing them to `formatPercentage()` to report `4.31%`.
- What should happen when two or more conditions have the same minimum or maximum value?
- How should conditions with no observed cases be handled?

To start with do this with only database 1:

```{r}
xc <- resultsCharacterisation |>
  filterResult(
    cohort_name %in% neuroCond,
    cdm_name == "Database 1",
    variable_name %in% c("Age", "Sex"),
    estimate_name %in% c("q25", "median", "q75", "percentage")
  ) |>
  tidy()
xp <- resultsPrevalence |>
  mutate(variable_name = "prevalence") |>
  filterResult(
    outcome_cohort_name %in% neuroCond,
    cdm_name == "Database 1"
  ) |>
  tidy() |>
  select(
    "cdm_name", "outcome_cohort_name", "variable_name", "outcome_count", 
    "denominator_count", "prevalence", "prevalence_95CI_lower",
    "prevalence_95CI_upper"
  )
```

To help you get started, the first sentence could be produced like this:

```{r}
numObs <- unique(xp$denominator_count)
```

The total number of individuals under observation was `r formatCount(numObs)`.

...


## Simple Table

Now we will create a table showing the prevalence of the conditions of interest in each database. First filter to the relevant conditions, then use `tablePrevalence()`. Populate its arguments so that the table is clear and easy to read.

```{r}
resultsPrevalence |>
  filterResult() |> # populate the filter accordingly
  tablePrevalence( # populate header, groupColumn, hide to make the table as you wish
    type = "flextable",
    style = "_brand.yml"
  )
```

For this prevalence table, we must explicitly specify `style` and `type`. This is not required for the other packages because they automatically detect these settings from the global options. In the future, `tablePrevalence()` may also support this behaviour.

We can add a caption to the table using a custom style defined in the Word reference document, as shown below:

:::{custom-style="TableCaption"}
Table 1. This is the title for table 1.
:::

```{r}
emptyTable()
```

:::{custom-style="CaptionFooter"}
IQR = interquartile range
:::

Use this example to create your own table with caption.

## Simple Figure

Similarly, create a figure showing the prevalence of the same conditions of interest using `plotPrevalence()`. Customise the plot so that it is clear and easy to read.

```{r}
resultsPrevalence |>
  filterResult() |> # populate the filter accordingly
  plotPrevalence( # populate x, y, colour, facet, ... to make the table as you wish
  )
```

## Loop of Sections

Now we will repeat the paragraph, figure, and table in a loop for every section. Producing multiple outputs inside a loop requires a slightly different approach. First, set the chunk option `results = "asis"` so that text generated with `cat()` is interpreted as Markdown. You will also need the helper functions defined in `functions.R`.

Objects returned at the top level of a chunk are printed automatically, but objects created inside a `for` loop are not. Therefore, tables and figures must be printed explicitly. `printTable()` converts and prints each table in a form that knitr can include, while `printFigure()` saves each plot and writes the corresponding image link into the document.

Let's see how it will look:

```{r, results='asis'}
sections <- unique(dic$section)
for (sec in sections) {
  # conditions of interest
  conditions <- dic$report_name[dic$section == sec]
  
  # to create the section title
  cat(paste0("### ", sec, "\n\n"))
  
  # data of interest
  xc <- resultsCharacterisation |>
    filterResult(
      cohort_name %in% conditions,
      variable_name %in% c("Age", "Sex"),
      estimate_name %in% c("q25", "median", "q75", "percentage")
    ) |>
    tidy()
  xp <- resultsPrevalence |>
    mutate(variable_name = "prevalence") |>
    filterResult(outcome_cohort_name %in% conditions) |>
    tidy() |>
    select(
      "cdm_name", "outcome_cohort_name", "variable_name", "outcome_count", 
      "denominator_count", "prevalence", "prevalence_95CI_lower",
      "prevalence_95CI_upper"
    )
  
  # Try to fit this text:
  # The total number of individuals under observation was XXX.
  # The neurological condition with the highest prevalence was XXX, with a
  # prevalence of XX% (95% CI: XX–XX%), while the condition with the lowest
  # prevalence was XXXX, with a prevalence of XX% (95% CI: XX–XX%).
  # The median age at diagnosis ranged from XX years (IQR: XX–XX) for XXXX
  # condition to XX years (IQR: XX–XX) for XXXX condition.
  # The proportion of females ranged from XX% in XXXX condition to XX% in XXXX
  # condition.
  
  # For example:
  formatProportion <- function(x) {
    sprintf("%.2f%% for %s in %s", x$percentage, x$cohort_name, x$cdm_name)
  }
  xfm <- xc |>
    filter(variable_level == "Female") |>
    filter(percentage == min(percentage))
  xfM <- xc |>
    filter(variable_level == "Female") |>
    filter(percentage == max(percentage))
  propFem <- paste0(
    "The proportion of females ranged from ",
    formatProportion(xfm),
    " to ",
    formatProportion(xfM),
    ".\n\n"
  )
  cat(propFem)
  
  # create tables
  captionTable("Table XX. Title")
  
  emptyTable() |>
    printTable() # note that now explicit printing is needed
  
  footer("Footer")
  
  # create figures
  
  emptyPlot() |>
    printFigure() # note that now explicit printing is needed
  
  captionFigure("Figure XX. Title")
  
  footer("Footer")
  
}
```

When you generalise the code, several edge cases will arise: sections containing only one condition, conditions with only males in the denominator, conditions that are absent from a database, and counts below five. Consider how your code should handle each case without failing or producing misleading text.

## Spacing in Quarto

You may have noticed that a single newline in the source does not create a line break in the rendered document. Markdown treats consecutive lines as part of the same paragraph.

To start a new paragraph, leave a blank line between blocks of text.

You can also insert a page break using `{{< pagebreak >}}`:

{{< pagebreak >}}

Adding several blank lines does not create extra vertical space: one or more blank lines simply separate paragraphs.






If you want to force a line break without starting a new paragraph, end the line with `\`, as shown below.

This is the first line. \
This starts a new line within the same paragraph.

If you want to add a visually empty line, place `&nbsp;` on its own line. Unlike additional blank lines in the source, it creates a paragraph containing a non-breaking space, which is preserved in the rendered document. The empty line before the next section is created in this way.

&nbsp;

## Change the style

- Table and figure styling is controlled by `_brand.yml`. Try changing the `darwin_blue` colour to `'#AF2F3E'`, changing the white lines to black, or experimenting with another setting. Regenerate the document to confirm that your changes have been applied.

- Caption and text styling is controlled by `data/template.docx`. In the Word styles pane, you will find the custom styles *TableCaption*, *FigureCaption*, and *CaptionFooter*. Try making *FigureCaption* bold or changing its colour, then regenerate the document to confirm that the changes have taken effect.

- Use the formatting functions defined earlier to control the number of decimal places displayed for percentages.

- Use the `options` argument to control the number of decimal places displayed in tables.

- Use `#` characters to create headings. The number of characters determines the heading level; for example, `####` creates a level-four heading and uses the *Heading 4* style from the Word reference document.
