---
title: "My report"
format:
  docx:
    fig-cap-location: top
    reference-doc: data/template.docx
execute:
  echo: false # R code is not printed
  message: false # R messages are not printed
  warning: false # R warning messages are not printed
---

```{r, eval=FALSE}
# make sure you have the latest version of the packages
#install.packages("pak")
pak::pkg_install(c(
  "visOmopResults", "omopgenerics", "IncidencePrevalence",
  "CohortCharacteristics", "here", "readr", "dplyr"
))
```

```{r}
# Check needed packages ----
if (packageVersion("visOmopResults") < "1.5.1") {
  stop("visOmopResults >= 1.5.1 is required")
}
if (packageVersion("omopgenerics") < "1.4.2") {
  stop("omopgenerics >= 1.4.2 is required")
}

# Load necessary packages ----
library(visOmopResults)
library(here)
library(readr)
library(omopgenerics)
library(dplyr)
library(IncidencePrevalence)
library(CohortCharacteristics)

# Load results ----
results <- importSummarisedResult(path = here("rawData"))

# source functions
source(file.path(getwd(), "functions.R"))

# Global options ----
knitr::opts_chunk$set(
  out.width  = "95%",  # figures occupy ~95% of document width
  out.height = "auto",
  dpi        = 320,    # ensure figure quality
  fig.width  = 6,      # default aspect ratio (can be overridden per-figure)
  fig.height = 3,
  results    = "asis"  # enable Markdown produced via cat() inside chunks
)

setGlobalPlotOptions(type = "ggplot") # the best configuration for word documents
setGlobalTableOptions(type = "flextable") # the best configuration for word documents
```

```{r}
# read the new names and sections for the report
dic <- read_csv("data/dictionary.csv", show_col_types = FALSE)
```

```{r}
# you can define some functions here so you can reuse them
formatCount <- function(x) {
  format(round(x), big.mark = ",")
}
formatPercentage <- function(x, n = 2) { # by default formated to 2 decimals
  sprintf(paste0("%.", n, "f%%"), x)
}
```

First, we will update the cohort names using the supplied dictionary. In the characterisation results, the cohort name is stored in the group field `cohort_name`, whereas in the prevalence results it is stored in `outcome_cohort_name`. We will therefore split the results into characterisation and prevalence results before renaming the cohorts.

```{r}
resultsCharacterisation <- results |>
  filterResult(result_type == "summarise_characteristics") |>
  splitGroup() |>
  left_join(
    dic |>
      select("cohort_name" = "original_name", "new" = "report_name"),
    by = c("cohort_name")
  ) |>
  select(!"cohort_name") |>
  rename("cohort_name" = "new") |>
  uniteGroup(cols = "cohort_name") |>
  newSummarisedResult()

resultsPrevalence <- results |>
  filterResult(result_type == "prevalence") |>
  splitGroup() |>
  left_join(
    dic |>
      select("outcome_cohort_name" = "original_name", "new" = "report_name"),
    by = c("outcome_cohort_name")
  ) |>
  select(!"outcome_cohort_name") |>
  rename("outcome_cohort_name" = "new") |>
  uniteGroup(cols = c("denominator_cohort_name", "outcome_cohort_name")) |>
  newSummarisedResult()
```

Does the code make sense?

In this first block of exercises we will work only with the neurological conditions.

```{r}
neuroCond <- dic |>
  filter(section == "Neurological conditions") |>
  pull("report_name")
```

You can use the variable that we have just created to print the neurological conditions programmatically. There are `r length(neuroCond)` neurological conditions: `r neuroCond`.

## Simple paragraph

Let's make our PI happy. We have agreed to include the following text:

The total number of individuals under observation was XXX.

The neurological condition with the highest prevalence was XXX, with a prevalence of XX% (95% CI: XX–XX%), while the condition with the lowest prevalence was XXXX, with a prevalence of XX% (95% CI: XX–XX%).

The median age at diagnosis ranged from XX years (IQR: XX–XX) for XXXX condition to XX years (IQR: XX–XX) for XXXX condition.

The proportion of females ranged from XX% in XXXX condition to XX% in XXXX condition.

When generating the text programmatically, consider:

- How many decimal places should be reported? You can use/define functions in the block that we created above.
- Prevalence and its confidence limits are stored as proportions (for example, `0.0431`), so multiply them by 100 before passing them to `formatPercentage()` to report `4.31%`.
- What should happen when two or more conditions have the same minimum or maximum value?
- How should conditions with no observed cases be handled?

To start with do this with only database 1:

```{r}
xc <- resultsCharacterisation |>
  filterResult(
    cohort_name %in% neuroCond,
    cdm_name == "Database 1",
    variable_name %in% c("Age", "Sex"),
    estimate_name %in% c("q25", "median", "q75", "percentage")
  ) |>
  tidy()
xp <- resultsPrevalence |>
  mutate(variable_name = "prevalence") |>
  filterResult(
    outcome_cohort_name %in% neuroCond,
    cdm_name == "Database 1"
  ) |>
  tidy() |>
  select(
    "cdm_name", "outcome_cohort_name", "variable_name", "outcome_count", 
    "denominator_count", "prevalence", "prevalence_95CI_lower",
    "prevalence_95CI_upper"
  )
```

To help you get started, the first sentence could be produced like this:

```{r}
numObs <- unique(xp$denominator_count)
```

The total number of individuals under observation was `r formatCount(numObs)`.

...


## Simple Table

Now we will create a table showing the prevalence of the conditions of interest in each database. First filter to the relevant conditions, then use `tablePrevalence()`. Populate its arguments so that the table is clear and easy to read.

```{r}
resultsPrevalence |>
  filterResult() |> # populate the filter accordingly
  tablePrevalence( # populate header, groupColumn, hide to make the table as you wish
    type = "flextable",
    style = "_brand.yml"
  )
```

Note that for this prevalence table we had to explicitly state the style and type, this is not needed for the rest of the packages as they automatically detect the style and type from global options. In the future (hopefully) this will be implemented and you won't need to manually state the style and type.

Now we want to add a caption to this table to do so we can use the predefined style template, see the example below:

:::{custom-style="TableCaption"}
Table 1. This is the title for table 1.
:::

```{r}
emptyTable()
```

:::{custom-style="CaptionFooter"}
IQR = interquartile range
:::

Use this example to create your own table with caption.

## Simple Figure

Similarly, create a figure showing the prevalence of the same conditions of interest using `plotPrevalence()`. Customise the plot so that it is clear and easy to read.

```{r}
resultsPrevalence |>
  filterResult() |> # populate the filter accordingly
  plotPrevalence( # populate x, y, colour, facet, ... to make the table as you wish
  )
```

## Loop of Sections

Now we will repeat the paragraph, figure, and table in a loop for every section. Producing multiple outputs inside a loop requires a slightly different approach. First, set the chunk option `results = "asis"` so that text generated with `cat()` is interpreted as Markdown. You will also need the helper functions defined in `functions.R`.

Objects returned at the top level of a chunk are printed automatically, but objects created inside a `for` loop are not. Therefore, tables and figures must be printed explicitly. `printTable()` converts and prints each table in a form that knitr can include, while `printFigure()` saves each plot and writes the corresponding image link into the document.

Let's see how it will look:

```{r, results='asis'}
sections <- unique(dic$section)
for (sec in sections) {
  # conditions of interest
  conditions <- dic$report_name[dic$section == sec]
  
  # to create the section title
  cat(paste0("### ", sec, "\n\n"))
  
  # data of interest
  xc <- resultsCharacterisation |>
    filterResult(
      cohort_name %in% conditions,
      variable_name %in% c("Age", "Sex"),
      estimate_name %in% c("q25", "median", "q75", "percentage")
    ) |>
    tidy()
  xp <- resultsPrevalence |>
    mutate(variable_name = "prevalence") |>
    filterResult(outcome_cohort_name %in% conditions) |>
    tidy() |>
    select(
      "cdm_name", "outcome_cohort_name", "variable_name", "outcome_count", 
      "denominator_count", "prevalence", "prevalence_95CI_lower",
      "prevalence_95CI_upper"
    )
  
  # Try to fit this text:
  # The total number of individuals under observation was XXX.
  # The neurological condition with the highest prevalence was XXX, with a
  # prevalence of XX% (95% CI: XX–XX%), while the condition with the lowest
  # prevalence was XXXX, with a prevalence of XX% (95% CI: XX–XX%).
  # The median age at diagnosis ranged from XX years (IQR: XX–XX) for XXXX
  # condition to XX years (IQR: XX–XX) for XXXX condition.
  # The proportion of females ranged from XX% in XXXX condition to XX% in XXXX
  # condition.
  
  # For example:
  formatProportion <- function(x) {
    sprintf("%.2f%% for %s in %s", x$percentage, x$cohort_name, x$cdm_name)
  }
  xfm <- xc |>
    filter(variable_level == "Female") |>
    filter(percentage == min(percentage))
  xfM <- xc |>
    filter(variable_level == "Female") |>
    filter(percentage == max(percentage))
  propFem <- paste0(
    "The proportion of females ranged from ",
    formatProportion(xfm),
    " to ",
    formatProportion(xfM),
    ".\n\n"
  )
  cat(propFem)
  
  # create tables
  captionTable("Table XX. Title")
  
  emptyTable() |>
    printTable() # note that now explicit printing is needed
  
  footer("Footer")
  
  # create figures
  
  emptyPlot() |>
    printFigure() # note that now explicit printing is needed
  
  captionFigure("Figure XX. Title")
  
  footer("Footer")
  
}
```

When you generalise the code, several edge cases will arise: sections containing only one condition, conditions with only males in the denominator, conditions that are absent from a database, and counts below five. Consider how your code should handle each case without failing or producing misleading text.

## Spacing in quarto

Probably
you
have
noticed
that adding just one
line break
does
not add the line break in the rendered document. If you want to 
create a new line you need

two line breaks.

Also you can add `{{< pagebreak >}}` if you want to add

{{< pagebreak >}}

a page break. As you can see multiple line






breaks do not add extra line breaks (e.g. space/1 line break -> space; 2 or more line breaks -> 1 line break).

If you explicitly want to add a line break you must use `\` see example:

\

This adds an extra line break.

## Change the style

- The style of tables and figures is controlled by the `_brand.yml` file change the darwin blue definition to: '#AF2F3E' or change the white lines into black or play with it. Regenerate the document so you can check that the changes have been applied properly. 

- The style of captions and text is controlled by the `data/template.docx` see on the styles part that *TableCaption*, *FigureCaption*, *CaptionFooter* have been defined. Play with them and add for example bold to the FigureCaption or change the colour or whatever you like. Regenerate the document so you see that the changes have been affected.

- Use the functions that you defined before to define the number of decimals used to display a percentage.

- Change the number of decimals used in the tables, you can control that using the `options` argument.

- For titles of different levels use '#' the more that you use will determine which is the level of the style used: e.g. if you use #### (4 hashtags) this will match the Heading 4 style.
